# Dictionary of geographical entities {#c22_Dicogeo}

```{r}
library(knitr)
library(dplyr)
library(WikidataR)
library(quanteda)
```





## Dictionary of geographical entities

We realize a test of the method described above on a preliminary dictionary of world regions combined with world states described by their names and capital cities


### Load the list of world regions, organisation states and capital cities

We start from a provisional list of 65 world regions based on continents (cont), other land area (14), sea regions (sea), regional organisations (org), states names (sta) and capital cities of states (cap).

```{r}
ent<-read.table("dict/wikigeo_V1.csv", 
                sep=";",
                header=T,
                encoding = "UTF-8")
table(ent$type)
```


### Extract définitions

We extract the definitions of the regions in the different languages with the function *extract_def()*

```{r, echo=FALSE}

extract_def <- function(item = c("Q15", "Q246"),
                        langs = c("fr","de","en","tr","ar")) {
  # Create empty dataset
  res<-data.frame()
  res$id    <- as.character()
  res$lang  <- as.character()
  res$label <- as.character()
  res$desc  <- as.character()
  
  
  # Loop of items
  n <- length(item)
  for (i in 1:n) {
    
     # Extract item properties
    item_prop <- get_property(item[i])[[1]]
  
   
     # Loop  of language
     p<-length(langs)
     for (j in 1:p) {
        id <- item[i]
        lang  <- langs[j]
        if(is.null(item_prop[["labels"]][[lang]]$value)==F) {label <- item_prop[["labels"]][[lang]]$value}
           else { label <- NA}
        if(is.null(item_prop[["descriptions"]][[lang]]$value)==F) {desc <- item_prop[["descriptions"]][[lang]]$value}
           else { desc <- NA}
        add <-data.frame(id,lang,label,desc)
        res<- rbind(res,add) 
        }
  
  }
  # Export result
return(res)

}
extract_alias <- function(items = c("Q15", "Q258"),
                          langs = c("fr","de","en","tr","ar")) {
  # Create empty dataset
  res<-data.frame()
  res$id    <- as.character()
  res$lang  <- as.character()
  res$label <- as.character()

  
  # Loop of items
  n <- length(items)
  for (i in 1:n) {
    
     # Extract item properties
    item_prop <- get_property(items[i])[[1]]
  
   
     # Loop  of language
     p<-length(langs)
     for (j in 1:p) {
        id <- items[i]
        lang  <- langs[j]
        if(is.null(item_prop[["labels"]][[lang]]$value)==F) {label <- item_prop[["labels"]][[lang]]$value} else { label <- NA}
        if(is.null(item_prop[["descriptions"]][[lang]]$value)==F) {desc <- item_prop[["descriptions"]][[lang]]$value}else { desc <- NA}
        add <-data.frame(id,lang,label)
        res<- rbind(res,add) 
           # Loop of aliases
              if (is.null(item_prop[["aliases"]][[lang]])==F) {
                ali <- item_prop[["aliases"]][[lang]]$value
                n<-length(ali)
               for (k in 1:n) {
                        label <- ali[k]
                        add <-data.frame(id,lang,label)
                        res<- rbind(res,add) 
                  }
              }
        
        }
  
  }
  # Export result
return(res)

}
```


```{r, eval=FALSE}
## NOT RUN : need several minutes !!!##
worldgeo_def <- extract_def(ent$wikidata,c("fr","de","en","tr","ar"))
names(ent)<-c("id","type_ent")
worldgeo_def<-worldreg_def %>% left_join(ent)
write.table(x = worldgeo_def,
             row.names = FALSE,
              file = "dict/worldgeo_def.csv",
              fileEncoding = "UTF-8",
              sep = ";")
saveRDS(object = worldgeo_def, file = "dict/worldgeo_def.RDS")

```

The dictionary of entities is further analyzed manually and completed by an internal code of entities specific to IMAGEUN's project




### Extract aliases and create dictionary

```{r, eval=FALSE}
## NOT RUN : very long time ...###
worldgeo_def <- readRDS("dict/worldgeo_def.RDS")

worldgeo_dict <- extract_alias(worldgeo_def$id, c("fr","de","en","tr","ar"))

write.table(x = worldgeo_dict,
              row.names = FALSE,
              file = "dict/worldgeo_dict.csv",
              fileEncoding = "UTF-8",
              sep = ";")
saveRDS(object = worldgeo_dict, file = "dict/worldgeo_dict.RDS")

```




